---
title: "p8105_hw3_ag3911"
output: github_document
---

#Problem 1

##Read and clean BRFSS data
* format to use appropriate variable names
* focus on "overall health" topic
* include only responses from "excellent" to "poor"
* organize responses as factor levels from "excellent" to "poor"
```{r}
install.packages("devtools", force = TRUE)
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
library(tidyverse)
brfss <- brfss_smart2010 %>%
  janitor::clean_names() %>%
  select(year, locationdesc, topic, response, data_value) %>%
  filter(topic == "Overall Health") %>% 
  select(-topic) %>%
  rename("Overall Health Response" = response) %>%
  rename("Overall Health Value" = data_value) %>%
  separate(locationdesc, into = c("state", "county"), sep = " - ") %>% 
  mutate(`Overall Health Response` = as.factor(`Overall Health Response`)) 

#do i need to do the following to change the factor levels from words to numbers? 
  #OH_fac = gl(1, 2, 3, 4, 5) %>%
  #levels(OH_fac) = c("Poor", "Fair", "Good", "Very Good", "Excellent")
```

##Q1: In 2002, which states were observed at 7 locations?

```{r}
brfss %>%
  filter(year == '2002') %>%
  group_by(state) %>%
  summarize(n = n_distinct(county)) %>%
  filter(n == 7)
```
In 2002, Connecticut, Florida, and North Carolina were observed at 7 county locations. 

##Q2: Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}
brfss_plot = brfss %>%
  distinct(year, state, county) %>%      #to keep only unique/distinct rows 
  group_by(year, state) %>%
  summarize(number = n())
ggplot(data = brfss_plot, aes(x = year, y = number, color = state)) + geom_line() 

#still need to add a graph title 
```


##Q3: Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
brfss_02_06_10 = brfss %>%
  filter(year == '2002' | year == '2006' | year == '2010') %>%
  filter(state == "NY") %>%
  spread(key = "Overall Health Response", value = "Overall Health Value") %>% 
  select(year, county, Excellent) %>% 
  group_by(year) %>%
  summarize(mean_ex = mean(Excellent), 
            sd_ex = sd(Excellent))
brfss_02_06_10
```

##Q4: For each year and state, compute the average proportion in each response category (taking the average across locations in a state). 

```{r}
brfss_avg = brfss %>%
  spread(key = "Overall Health Response", value = "Overall Health Value") %>% 
  group_by(year, state) %>%
  summarize(mean_excellent = mean(Excellent), 
            mean_vgood = mean(`Very good`),
            mean_good = mean(Good),
            mean_fair = mean(Fair),
            mean_poor = mean(Poor))
brfss_avg
```

##Q5: Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss %>%
  #select(year, state, `Overall Health Response`) %>% 
  group_by(year, state, `Overall Health Response`) %>%
  summarize(mean = mean(`Overall Health Value`)) %>%
  ggplot(aes(x = year, y = mean, color = state)) +
  geom_line() + 
  facet_grid(~`Overall Health Response`)
```

#Problem 2

##Read and clean Instacart data
```{r}
#install.packages("devtools", force = TRUE)
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
library(tidyverse)

instacart = instacart %>%
  janitor::clean_names() %>%
  filter(aisle != 'missing')
```
The size of the dataset is `r dim(instacart)`. It contains information about orders, products ordered, and customers. Order information includes the order ID number, the order in which each product was added to the online shopping cart, whether or not reordering in the cart occurred, the day of the week and hour of the day it was placed. For example, an order placed on day of the week 5 at hour 17 means that it was placed on Friday at 5 pm. Product information includes the product ID number, product name, the name of the aisle in which its found, and the ID number for and name of the department in which it's found. For example, the dairy eggs department has ID number 16.  Customer information includes the customer ID number, their order sequence number, and the number of days since their last order. 

##Q1: How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
    group_by(aisle_id) %>%
    summarise(n = n())
#to double check, count aisles using name instead of id. should get the same number of aisles
instacart %>%      
    group_by(aisle) %>%
    summarize(n = n())
```
There are 133 aisles. 

```{r}
#to find from which aisles the most items are ordered
cart = instacart %>%
    group_by(aisle) %>%
    summarize(number = n_distinct(product_id)) %>%
    filter(min_rank(desc(number)) < 4)
```
The most items are ordered from 1) the "candy chocolate" aisle followed by 2) the "yogurt" aisle followed by 3) the "ice cream ice" aisle.  

##Q2: Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
    group_by(aisle) %>%
    summarize(number = n_distinct(product_id)) %>%
    #filter(rank(desc(number))) %>%
    ggplot(aes(x = aisle , y = number))
```

##Q3: Make a table showing the most popular item in aisles “baking ingredients”, in “dog food care”, and in “packaged vegetables fruits”
```{r}
cart_bake_dog_vegf = instacart %>%
  filter(aisle == 'baking ingredients' | aisle == 'dog food care' | aisle == 'packaged vegetables fruits') %>%
  select(product_name, aisle) %>% 
  group_by(aisle) %>%
  summarise(n_product = n_distinct(product_name))

  summarize(max = max(), na.rm = FALSE) 

##Q4: Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


#Problem 3

##Read and Clean New York NOAA data 
*Create separate variables for year, month, and day. 
*Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 
*For snowfall, what are the most commonly observed values? Why?
```{r}
#install.packages("devtools", force = TRUE)
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
library(tidyverse)

ny_noaa %>% 
  janitor::clean_names()
```

##write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following questions (commenting on the results of each):



##Q1: Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
```

##Q2: Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
```

